Joint training of classification model and similarity model for low-resource text classification in chatbot

### 2020-08-23 chatbot_classification/aaai_2021/paper.ver_02.tex
version2.

One of the most popular approach among existing work was text classification model based on neural network. Since the length of the text is quite short, many previous work\cite{wen2016network} choose neural network such as convolutional neural networks (CNNs)\cite{kim2014convolutional,zhang2015character,conneau2016very}
or long short term memory networks
(LSTMs)\cite{mousa2017contextual,liu2016recurrent} to accomplish the task of extracting semantic feature from limited amount of words. A common model structure is adding a softmax classifier to the top of the neural network. Another common approach is text classification neural network based on label-word joint embedding, such as LEAM\cite{wang2018joint}. This approach involves label embedding when constructing the text-sequence representation, which poses requirements on label information. In our cases, since we have large amount of class labels without being well-defined, this approach can do only a few help to the low-resource chatbot scenario. Afterwards, pre-trained language models on large corpus like BERT\cite{devlin2018bert} and
RoBERTa\cite{liu2019roberta} has  been proven more powerful in solving many NLP
tasks      including      short-text      classification\cite{madabushi2020cost}.   Especially   for  few-shot  scenarios\cite{yu2018diverse}, pre-trained  model  based on transformers\cite{vaswani2017attention}, shown in Fig. \ref{Fig.1.sub.1}, tends  to  do more help to reducing the negative effects brought by scarcity of training data.

Another popular approach among previous work was based on sentence-pair model. The motivation of this approach was started from the idea of Information Retrival (IR) based chatbot\cite{jafarpour2010filter, leuski2011npceditor}. Having a Q-A (Question-Answer) pairs dataset and user query Q, the IR based conversational system will look up in the Q-A dataset for the pair (Q', A') that best matches query Q through semantic analysis and returns A' as the answer to Q\cite{mnasri2019recent}. In this way, sentence-pair model pre-trained on large corpus of semantic similarity identification task can be applied for being used as a tool to identify the class with highest semantic similarity to customer's query. A common model structure of pre-trained\cite{devlin2018bert} sentence-pair model is based on multiple cross-attention mechanism\cite{barkan2020scalable}, shown in Fig. \ref{Fig.1.sub.2}. Due to the fact that many experiment results have shown that RoBERTa\cite{liu2019roberta} is an improved  version  of  BERT, and has achieved amazing results on both text classification and sentence-pairs semantic similarity tasks, we choose RoBERTa as an important baseline approach in this paper.

Despite the success of these 2 approaches, they still have some limitations in task-specific chatbot scenario. As for the text classification model approach, it's quite hard to use data augmentation method to solve the data insufficiency challenge since the it's usually unfeasible to get large amount of domain-specific data when facing a new task. With respect to the sentence-pair model approach, though we can obtain large amount of data in semantically duplicate sentence pair identification domain for transfer learning\cite{sun2019fine}, it's still quite hard to perform well on task-specific dataset because the training objective of sentence-pair model is semantic similarity, which is slightly different from the target of classification task. That is to say, there always exists some intent classes which cannot be distinguished by each other merely depending on semantic similarity. For example, we have 2 user query saying, A: What should I do if I want to change my password? B: What is the modification rule if I want to change my password? In this case, sentence A is semantically similar to B, since they both express the 
desire for changing password. However, it's still reasonable to classify them into 2 intent classes, since A is asking for the method for changing password, while B is asking for the rule to follow when creating a new password.

@@

LEAM & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{$0.4545$} & \multirow{2}*{$0.5554$} & \multirow{2}*{$0.5936$} & \multirow{2}*{$0.6599$} & \multirow{2}*{$0.7086$} & \multirow{2}*{$0.6091$}\tabularnewline

@@

\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
\multicolumn{1}{|c|}{\multirow{2}*{Dataset}}&\multicolumn{1}{c|}{\multirow{2}*{Model}}& $K=3$ & $K=5$ & $K=10$ & $K=15$ & $K=20$\tabularnewline
\multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & $P=20$ & $P=10$ & $P=5$ & $P=4$ & $P=3$\tabularnewline
\hline
\multicolumn{1}{|c|}{\multirow{2}*{ITG}}& 2-stage SFC & $0.8034$ & $0.8124$ & $0.8008$ & $0.7967$ & $0.7934$\tabularnewline
\cline{2-7}
\multicolumn{1}{|c|}{} & Hierarchical SFC & $0.7882$ & $0.8114$ & $0.8010$ & $0.7972$ & $0.7826$\tabularnewline
\hline
\multicolumn{1}{|c|}{\multirow{2}*{Amazon-670k}}& 2-stage SFC & $0.7278$ & $0.7364$ & $0.7366$ & $0.7328$ & $0.7204$\tabularnewline
\cline{2-7}
\multicolumn{1}{|c|}{} & Hierarchical SFC & $0.7334$ & $0.7445$ & $0.7516$ & - & -\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\caption{We show the performance of SFC from different settings of hyperparameters for $K$, the candidate class number, and $P$, the sample number for each class.}

@@

\textbf{Comparison among SFC variants.} First, we turn to compare among the four SFC variants shown in Table \ref{tbe:table2} to analyze the improvement brought by multi-task and hierarchical joint-model structure. We conduct ablation experiments to study the benefits brought by multi-task training by comparing multi-task 2-stage SFC with single-task 2-stage SFC. We can see from Table \ref{tbe:table2} that 2-stage SFC with only task 1 (sentence-pair similarity task) degrades by 0.65 percent on average and 2-stage SFC with only task 2 (top-K classification task) degrades by 0.3 percent on average. The degradation indicates that training with these 2 tasks together in SFC can effectively help the sentence-pair model extract more knowledge from the limited amount of data. Besides, task 2 degrades 0.35 percent less than task 1. This observation suggests that task 2 may play a relatively more important role in multi-task learning, which leads us to put larger weight $\alpha_{K}$ to the loss $\mathcal{L}^{K}$ of task 2 when calculating the total loss $\mathcal{L}$. Also, we can see that the hierachical SFC steadily outperforms 2-stage SFC on 5 diverse datasets by 2 percents on average. This observation supports the idea that the hierachical joint model structure can alleviate the limitation brought by the lack of interaction between similarity model and classification model in 2-stage SFC. Therefore, for the following analysis, we will focus on the comparison between hierachical SFC and the other baseline models.

\textbf{Analysis under few-shot scenarios.} Next, we analyze the overall improvement trend of SFC under various few-shot settings. Our main focus is to study chatbot building in common situation during real-world application where only a few sample sentences are available for each class. The experimental results for the 3 diverse intent detection datasets under few-shot settings indicate that our proposed SFC can achieve +2.47 percents average improvement over one of the most powerful baseline models, which is RoBERTa-large classification model, on all these few-shot settings. Moreover, the improvement in F-score becomes more and more prominent with the decrease of data sample number per class. Our hierarchical SFC can achieve +4.97 percents average improvements over RoBERTa-large classification model in the most extreme few-shot setting with only 5 training samples per class. This observation indicates that SFC is especially suitable for being applied during the initial stage of building task-specific chatbot system, where the amount of data sample for each class label is extremely scarce.

\textbf{Fusion of classification and similarity model.} Then, we begin to compare the general performance of SFC with single-task classification and similarity baselines. As for the classification model, hierarchical SFC outperforms TextCNN by +37.99 percents, ALBERT-xxlarge by +17.87 percents\footnote{We found that the performance of ALBERT.xxlarge is extremely unstable for classification task. Here, we did multiple run with different settings of parameters such as learning rate and batch size for each dataset, and record the best result we can obtain in Table \ref{tbe:table2}.}, BERT-large by +9.81 percents, RoBERTa-base by +4.52 percents, RoBERTa-large by +2.47 percents on average F-score over 5 datasets. In this way, we can see that RoBERTa is the best baseline pre-trained model for the low-resource text classification task we are studying. As for the similarity model, we only choose the strongest pretrained model RoBERTa-large as our baseline, and hierarchical SFC achieves the improvement of +7.81 percents on average. Moreover, hierarchical SFC's result on 2 datasets without data sampling for few-shot settings, which are ITG and Amazon-670k, are also 1.24 and 2.89 percents higher than Roberta-large classification model, and are 6.96 and 10.77 percents higher than Roberta-large similarity model. These improvements illustrate that putting classification model and similarity model at different network layers to cooperate with each other is a successful design of hierarchical network structure for joint model, which enables these two models assist each other for better overall results.

>>>>

One of the most popular approach among existing work was text classification model based on neural network. Since the length of the text is quite short, many previous work\cite{wen2016network} choose neural network such as convolutional neural networks (CNNs)\cite{kim2014convolutional,zhang2015character,conneau2016very}
or long short term memory networks
(LSTMs)\cite{mousa2017contextual,liu2016recurrent} to accomplish the task of extracting semantic feature from limited amount of words. A common model structure is adding a softmax classifier to the top of the neural network. Afterwards, pre-trained language models on large corpus like BERT\cite{devlin2018bert} and
RoBERTa\cite{liu2019roberta} has  been proven more powerful in solving many NLP
tasks      including      short-text      classification\cite{madabushi2020cost}.   Especially   for  few-shot  scenarios\cite{yu2018diverse}, pre-trained  model  based on transformers\cite{vaswani2017attention}, shown in Fig. \ref{Fig.1.sub.1}, tends  to  do more help to reducing the negative effects brought by scarcity of training data.

Another common approach is text classification neural network based on label-word joint embedding, such as LEAM\cite{wang2018joint}, MTLE\cite{zhang2017multi} and EXAM\cite{du2019explicit}. This approach involves label embedding when constructing the text-sequence representation. Take LEAM as an example, the text classification model is implemented by jointly embedding the word and label in the same latent space\cite{wang2018joint}. Afterwards, the text representation are constructed based on text-label compatibility through attention mechanism. In this way, additional sources of semantic information from class labels can be fully leveraged, which is quite helpful especially under few-shot scenario.

The third popular approach among previous work was based on sentence-pair model. The motivation of this approach was started from the idea of Information Retrival (IR) based chatbot\cite{jafarpour2010filter, leuski2011npceditor}. Having a Q-A (Question-Answer) pairs dataset and user query Q, the IR based conversational system will look up in the Q-A dataset for the pair (Q', A') that best matches query Q through semantic analysis and returns A' as the answer to Q\cite{mnasri2019recent}. In this way, sentence-pair model pre-trained on large corpus of semantic similarity identification task can be applied for being used as a tool to identify the class with highest semantic similarity to customer's query. A common model structure of pre-trained\cite{devlin2018bert} sentence-pair model is based on multiple cross-attention mechanism\cite{barkan2020scalable}, shown in Fig. \ref{Fig.1.sub.2}. Due to the fact that many experiment results have shown that RoBERTa\cite{liu2019roberta} is an improved  version  of  BERT, and has achieved amazing results on both text classification and sentence-pairs semantic similarity tasks, we choose RoBERTa as an important baseline approach in this paper.

Despite the success of these 3 approaches, they still have some limitations in task-specific chatbot scenario. As for the text classification model approach, it's quite hard to use data augmentation method to solve the data insufficiency challenge since the it's usually unfeasible to get large amount of domain-specific data when facing a new task. As for label-word joint embedding approach, it poses noticeable requirements on label information. However, in our cases, since we have large amount of class labels without being well-defined for question answering chatbot, the only thing we can do is to choose one of the user utterance as the class label to represent the key information of the whole class. In this way, the label will become not only prolix but also biased, which will bring some negative effect to label embedding. With respect to the sentence-pair model approach, though we can obtain large amount of data in semantically duplicate sentence pair identification domain for transfer learning\cite{sun2019fine}, it's still quite hard to perform well on task-specific dataset because the training objective of sentence-pair model is semantic similarity, which is slightly different from the target of classification task. That is to say, there always exists some intent classes which cannot be distinguished by each other merely depending on semantic similarity. For example, we have 2 user query saying, A: What should I do if I want to change my password? B: What is the modification rule if I want to change my password? In this case, sentence A is semantically similar to B, since they both express the 
desire for changing password. However, it's still reasonable to classify them into 2 intent classes, since A is asking for the method for changing password, while B is asking for the rule to follow when creating a new password.

@@

LEAM & \multirow{2}*{$0.7514$} & \multirow{2}*{$0.8203$} & \multirow{2}*{$0.8612$} & \multirow{2}*{$0.5422$} & \multirow{2}*{$0.7812$} & \multirow{2}*{$0.8129$} & \multirow{2}*{$0.8280$} & \multirow{2}*{$0.4545$} & \multirow{2}*{$0.5554$} & \multirow{2}*{$0.5936$} & \multirow{2}*{$0.6599$} & \multirow{2}*{$0.7086$} & \multirow{2}*{$0.6091$}\tabularnewline

@@

\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
\multicolumn{1}{|c|}{\multirow{2}*{Dataset}}&\multicolumn{1}{c|}{\multirow{2}*{Model}}& $K=3$ & $K=5$ & $K=10$ & $K=15$ & $K=20$\tabularnewline
\multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & $P=20$ & $P=10$ & $P=5$ & $P=4$ & $P=3$\tabularnewline
\hline
\multicolumn{1}{|c|}{\multirow{2}*{ITG}}& 2-stage SFC & $0.8034$ & $0.8124$ & $0.8008$ & $0.7967$ & $0.7934$\tabularnewline
\cline{2-7}
\multicolumn{1}{|c|}{} & Hierarchical SFC & $0.7882$ & $0.8114$ & $0.8010$ & $0.7972$ & $0.7826$\tabularnewline
\hline
\multicolumn{1}{|c|}{\multirow{2}*{Amazon-670k}}& 2-stage SFC & $0.7278$ & $0.7364$ & $0.7366$ & $0.7328$ & $0.7204$\tabularnewline
\cline{2-7}
\multicolumn{1}{|c|}{} & Hierarchical SFC & $0.7334$ & $0.7445$ & $0.7516$ & $0.7344$ & $0.7373$\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\caption{We show the performance of SFC from different settings of hyperparameters for $K$, the candidate class number, and $P$, the sample number for each class.}

@@

\textbf{Comparison among SFC variants.} First, we turn to compare among the four SFC variants shown in Table \ref{tbe:table2} to analyze the improvement brought by multi-task and hierarchical joint-model structure. We conduct ablation experiments to study the benefits brought by multi-task training by comparing multi-task 2-stage SFC with single-task 2-stage SFC. We can see from Table \ref{tbe:table2} that 2-stage SFC with only task 1 (sentence-pair similarity task) degrades by 0.65 percent on average and 2-stage SFC with only task 2 (top-K classification task) degrades by 0.3 percent on average. The degradation indicates that training with these 2 tasks together in SFC can effectively help the sentence-pair model extract more knowledge from the limited amount of data. Besides, task 2 degrades 0.35 percent less than task 1. This observation suggests that task 2 may play a relatively more important role in multi-task learning, which leads us to put larger weight $\alpha_{K}$ to the loss $\mathcal{L}^{K}$ of task 2 when calculating the total loss $\mathcal{L}$. Also, we can see that the hierachical SFC steadily outperforms 2-stage SFC on 5 diverse datasets by 1.17 percents on average. This observation supports the idea that the hierachical joint model structure can alleviate the limitation brought by the lack of interaction between similarity model and classification model in 2-stage SFC. Therefore, for the following analysis, we will focus on the comparison between hierachical SFC and the other baseline models.

\textbf{Analysis under few-shot scenarios.} Next, we analyze the overall improvement trend of SFC under various few-shot settings. Our main focus is to study chatbot building in common situation during real-world application where only a few sample sentences are available for each class. The experimental results for the 3 diverse intent detection datasets (ClINC150, BANKING77, HWU64) under few-shot settings indicate that our proposed 
hierarchical SFC can achieve 2.47 percents average improvement over one of the most powerful baseline models, which is RoBERTa-large classification model, on all these few-shot settings. Moreover, the improvement in F-score becomes more and more prominent with the decrease of data sample number per class. Our hierarchical SFC can achieve 4.97 percents average improvements over RoBERTa-large classification model in the most extreme few-shot setting with only 5 training samples per class. This observation indicates that SFC is especially suitable for being applied during the initial stage of building task-specific chatbot system, where the amount of data sample for each class label is extremely scarce.

\textbf{Comparison between label embedding and sentence pair embedding.} Afterwards, we start to compare the performance of SFC with the text-label joint embedding baseline named as LEAM. As shown in Table \ref{tbe:table2}, hierarchical SFC outperforms LEAM by 14.25 percents on average F-score over 5 datasets. Besides, although the labels within the 3 diverse intent detection datasets (CLINC150, BANKING77, HWU64) are all being well-defined as the semantic summary of each intent, hierarchical SFC still outperforms LEAM by 14.69 percents on average F-score. This observation indicates that it's usually quite hard to use a fixed label with only a few words as the representative of all the sentences within a class since there are always plenty of ways for people to express questions or queries about the same topic during a conversation with chatbot. Therefore, learning the semantic similarity between user query and sample sentences through sentence pair embedding can be more reliable than making decision according to the semantic relationship between user query and class labels through label embedding.

\textbf{Fusion of classification and similarity model.} Then, we begin to compare the general performance of SFC with single-task classification and similarity baselines. As for the classification model, hierarchical SFC outperforms TextCNN by 26.19 percents, ALBERT-xxlarge by 17.87 percents\footnote{We found that the performance of ALBERT.xxlarge is extremely unstable for classification task. Here, we did multiple run with different settings of parameters such as learning rate and batch size for each dataset, and record the best result we can obtain in Table \ref{tbe:table2}.}, BERT-large by 9.81 percents, RoBERTa-base by 4.52 percents, RoBERTa-large by 2.47 percents on average F-score over 5 datasets. In this way, we can see that RoBERTa is the best baseline pre-trained model for the low-resource text classification task we are studying. As for the similarity model, we only choose the strongest pretrained model RoBERTa-large as our baseline, and hierarchical SFC achieves the improvement of 7.81 percents on average. Moreover, hierarchical SFC's result on 2 datasets without data sampling for few-shot settings, which are ITG and Amazon-670k, are also 1.24 and 2.89 percents higher than Roberta-large classification model, and are 6.96 and 10.77 percents higher than Roberta-large similarity model. These improvements illustrate that putting classification model and similarity model at different network layers to cooperate with each other is a successful design of hierarchical network structure for joint model, which enables these two models assist each other for better overall results.
